\bf An Unitary RG-based Auxiliary Model Approach\\ to Strongly-Correlated Systems

### 2023-02-02 notes/notes.tex
updated some older figures and removed unnecessary ones

\documentclass{report}
\usepackage{amsthm,amsmath,amssymb,braket,multirow,caption,relsize,tocbibind,color,appendix,authblk,enumitem}
% \usepackage[draft]{graphicx}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=0.5in,bottom=1in]{geometry}

@@

\title{A New Unitary RG-based Auxiliary Model Approach to Strongly-Correlated Systems}
\author{Abhirup Mukherjee, Siddhartha Lal}
\affil{Indian Institute of Science Education and Research Kolkata, Mohanpur}
\date{\today}

@@

\section{Evidence for the Mott MIT}
Through the URG study of the generalised SIAM with correlation (eq.~\ref{siam_attr}) shown in chapter \ref{siam-attr-chap}, we can conclude the following about that model:
\begin{itemize}
	\item For \(U_b > J/4\), the couplings \(J\) and \(V\) become irrelevant while \(U\) is relevant, which means that \textbf{the impurity gets cut off from the bath at low energies}.
\begin{center}
	\includegraphics[width=0.7\textwidth]{../figures/phase-map-MIT.pdf}
\end{center}
\item The impurity spectral function \(A_{dd}(\omega)\) and the impurity-bath real space off-diagonal Greens function \(A_{d0}(\omega)\) reveal a gap in the spectrum at low \(\omega\) for \(U_b > J/4\). The gap in the diagonal spectral function shows that the impurity site cannot fluctuate through gapless excitations, while the vanishing of the off-diagonal spectral function shows that an electron on the impurity site cannot delocalise out of that site.  Other measures like mutual information and spin-spin correlations also show a sharp change on crossing that value, demonstrating \textbf{the destruction of the Kondo cloud and the stabilisation of the local moment on the impurity}.

\begin{figure}[!htb]
\centering
	\includegraphics[width=0.3\textwidth]{../figures/spec_func_Ub_by_J=-0.100.pdf}
	\includegraphics[width=0.3\textwidth]{../figures/spec_func_Ub_by_J=-0.250.pdf}
	\includegraphics[width=0.3\textwidth]{../figures/spec_func_Ub_by_J=-0.26.pdf}
	\includegraphics[width=0.3\textwidth]{../figures/spec_func_d0_Ub_by_J=-0.100.pdf}
	\includegraphics[width=0.3\textwidth]{../figures/spec_func_d0_Ub_by_J=-0.250.pdf}
	\includegraphics[width=0.3\textwidth]{../figures/spec_func_d0_Ub_by_J=-0.26.pdf}
	\caption{Evolution of diagonal and off-diagonal spectral functions across the transition.}
	\label{spec_func_mit}
\end{figure}

\begin{figure}[!htb]
\centering
	\includegraphics[width=0.3\textwidth]{../figures/corr-k-t=1.000,J=10.000,0.000,30,V=3J,Ubath=-U_by_10,N=6,U=1.000,889.513,30.pdf}
	\includegraphics[width=0.3\textwidth]{../figures/corr-d0-t=1.000,J=10.000,0.000,30,V=3J,Ubath=-U_by_10,N=6,U=1.000,889.513,30.pdf}
	\includegraphics[width=0.3\textwidth]{../figures/r1p-t=1.000,J=10.000,0.000,30,V=3J,Ubath=-U_by_10,N=6,U=1.000,889.513,30.pdf}
	\caption{Evolution of correlation functions and mutual information across the transition.}
\end{figure}

@@

\begin{figure}[!htb]
\centering
	\includegraphics[width=0.75\textwidth]{../figures/overlaps_gs-J=10.000.pdf}
	\caption{Evolution of the zero-bandwidth ground state across the transition.}
\end{figure}

@@

On the other side \(r > r^{*}\) of the critical point, we know that the impurity gets decoupled from the bath, such that the weighted poles \(\left(|d^p_n|^2 + \left(d^p_n\right)^* z^p_n\right) \left(\mathcal{G}_\text{aux}(d, \omega)\right)_{nn}\) and \(\left(|d^h_n|^2 + \left(z^h_n\right)^* d^h_n\right)\left(\mathcal{G}_\text{aux}(d, -\omega)\right)_{nn}\) vanish. This is simply the loss of spectral weight at low-\(\omega\) because of the removal of the gapless excitations, and is demonstrated through both the diagonal and off-diagonal spectral functions in fig.~\ref{spec_func_mit}. This then implies that the bulk Greens functions will also gap out at low energies, and that describes an insulating phase where the electrons get "jammed" in the local states. By the same argument as above, the bulk spectral function also sees a gap in this phase when the impurity spectral function is gapped (see eq.\eqref{main}).

@@

\section{\(k-\)space picture of the MIT in terms of the momentum-resolved self-energy \(\Sigma(\vec k,\omega)\)}

@@

These plots are available as videos \href{https://abhirup-m.github.io/misc-sigma/}{\underline{here}}.

\clearpage
\section{Some important remarks}
\subsection{Entanglement as an order parameter for the transition}
The spectral weights in the diagonal Greens function can be written in terms of a geometric measure of entanglement:
\begin{equation}\begin{aligned}
	d_n^p = \braket{\Phi_n | c^\dagger_{d\sigma} | \Phi_0} = \braket{\Phi_n | c^\dagger_{d\sigma} \sum_n \ket{\phi_m} \bra{\phi_m} | \Phi_0}
\end{aligned}\end{equation}
where, \(\ket{\phi_m}\) are the eigenstates of the Hamiltonian composed of a two-site system (formed by the impurity site and the zeroth site) in direct product with the rest of the lattice sites. The two site system interacts only via a spin-exchange coupling \(J\), and the impurity and bath sites have local correlations \(U\) and \(U_b\) respectively. The ground state is the spin-singlet in direct product with the ground state of the tight-binding chain formed by the remaining sites. The summation involves inner products of the ground state \(\ket{\Phi_0}\) with the eigenstates \(\ket{\phi_m}\). Only two states out of all \(\ket{\phi_m}\) give non-zero overlap - the spin singlet state \(\ket{SS}\) and the charge triplet-zero state \(\ket{CT}\). These two non-zero overlaps give rise to non-zero entanglement: \(\varepsilon_\text{SS} = 1 - |\braket{\Phi_0 | SS}|^2, \varepsilon_\text{CT} = 1 - |\braket{\Phi_0 | CT}|^2\). The spectral weight becomes
\begin{equation}\begin{aligned}
	d_n^p = \bra{\Phi_n} | c^\dagger_{d\sigma} \left(\ket{SS} + \ket{CT} \right) \left(\sqrt{1 - \varepsilon_\text{SS}} + \sqrt{1 - \varepsilon_\text{CT}}\right)
\end{aligned}\end{equation}
The total spectral weight can be written as
\begin{equation}\begin{aligned}
	|d_n^p|^2 = |\bra{\Phi_n} | c^\dagger_{d\sigma} \left(\ket{SS} + \ket{CT} \right)|^2 \left(2 - \varepsilon_\text{SS} - \varepsilon_\text{CT} + 2\sqrt{1 - \varepsilon_\text{SS}}\sqrt{1 - \varepsilon_\text{CT}}\right)
\end{aligned}\end{equation}
Near the metal, the ground state is, to a very good approximation, the state \(\ket{SS}\). One can then say that the entanglement \(\varepsilon_\text{SS}\) is poor - the entangled state is very close to the "separable" state. The term separable is used here in the sense that the metallic state defines a limit of the system. As \(U_b\) increases, \(\varepsilon_\text{SS}\) increases. On the other hand, the entanglement with the charge triplet is maximum in the metal, because the ground state does not have much of the charge triplet. Towards the MIT, this contribution increases, and the entanglement decreases.

>>>>

\documentclass{report}
\usepackage{amsthm,amsmath,amssymb,braket,multirow,caption,relsize,tocbibind,color,appendix,authblk,enumitem}
% \usepackage[draft]{graphicx}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=0.7in,bottom=1in]{geometry}

@@

\title{\bf An Unitary RG-based Auxiliary Model Approach\\ to Strongly-Correlated Systems}
\author{Abhirup Mukherjee, Siddhartha Lal}
\affil{Indian Institute of Science Education and Research Kolkata, Mohanpur}
\date{\today}

@@

\section{Evidence for the Mott MIT}
Through the URG study of the generalised SIAM with correlation (eq.~\ref{siam_attr}) shown in chapter \ref{siam-attr-chap}, we can conclude the following about that model:
\begin{itemize}
	\item For \(U_b > J/4\), the couplings \(J\) and \(V\) become irrelevant while \(U\) is relevant, which means that \textbf{the impurity gets cut off from the bath at low energies}.
\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.48\textwidth]{../figures/phase-map-MIT.pdf}
	\includegraphics[width=0.48\textwidth]{../figures/spectral-function.pdf}
	\caption{{\it Left:} Phase diagram. {\it Right:} Impurity spectral function}
	\label{spec_func_mit}
\end{figure}

\item The impurity spectral function \(A_{dd}(\omega)\) and the impurity-bath real space off-diagonal Greens function \(A_{d0}(\omega)\) reveal a gap in the spectrum at low \(\omega\) for \(U_b > J/4\). The gap in the diagonal spectral function shows that the impurity site cannot fluctuate through gapless excitations, while the vanishing of the off-diagonal spectral function shows that an electron on the impurity site cannot delocalise out of that site.  Other measures like mutual information and spin-spin correlations also show a sharp change on crossing that value, demonstrating \textbf{the destruction of the Kondo cloud and the stabilisation of the local moment on the impurity}.

\begin{figure}[!htb]
\centering
	\includegraphics[width=0.48\textwidth]{../figures/corrs_gs.pdf}
	\includegraphics[width=0.48\textwidth]{../figures/spin-charge-corr-full.pdf}
	\includegraphics[width=0.48\textwidth]{../figures/mutinfo-d0-01-full.pdf}
	\includegraphics[width=0.48\textwidth]{../figures/doub_occ.pdf}
\end{figure}

@@

@@

On the other side \(r > r^{*}\) of the critical point, we know that the impurity gets decoupled from the bath, such that the weighted poles \(\left(|d^p_n|^2 + \left(d^p_n\right)^* z^p_n\right) \left(\mathcal{G}_\text{aux}(d, \omega)\right)_{nn}\) and \(\left(|d^h_n|^2 + \left(z^h_n\right)^* d^h_n\right)\left(\mathcal{G}_\text{aux}(d, -\omega)\right)_{nn}\) vanish. This is simply the loss of spectral weight at low-\(\omega\) because of the removal of the gapless excitations, and is demonstrated through the impurity spectral function in fig.~\ref{spec_func_mit}. This then implies that the bulk Greens functions will also gap out at low energies, and that describes an insulating phase where the electrons get "jammed" in the local states. By the same argument as above, the bulk spectral function also sees a gap in this phase when the impurity spectral function is gapped (see eq.\eqref{main}).

@@

\comment{
\section{\(k-\)space picture of the MIT in terms of the momentum-resolved self-energy \(\Sigma(\vec k,\omega)\)}

@@

These plots are available as videos \href{https://abhirup-m.github.io/misc-sigma/}{\underline{here}}.
}

\section{Some important remarks}
\subsection{Entanglement as an order parameter for the transition}
The spectral weights in the diagonal Greens function can be written in terms of a geometric measure of entanglement:
\begin{equation}\begin{aligned}
	d_n^p = \braket{\Phi_n | c^\dagger_{d\sigma} | \Phi_0} = \braket{\Phi_n | c^\dagger_{d\sigma} \sum_n \ket{\phi_m} \bra{\phi_m} | \Phi_0}
\end{aligned}\end{equation}
where, \(\ket{\phi_m}\) are the eigenstates of the Hamiltonian composed of a two-site system (formed by the impurity site and the zeroth site) in direct product with the rest of the lattice sites. The two site system interacts only via a spin-exchange coupling \(J\), and the impurity and bath sites have local correlations \(U\) and \(U_b\) respectively. The ground state is the spin-singlet in direct product with the ground state of the tight-binding chain formed by the remaining sites. The summation involves inner products of the ground state \(\ket{\Phi_0}\) with the eigenstates \(\ket{\phi_m}\). Only two states out of all \(\ket{\phi_m}\) give non-zero overlap - the spin singlet state \(\ket{SS}\) and the charge triplet-zero state \(\ket{CT}\). These two non-zero overlaps give rise to non-zero entanglement: \(\varepsilon_\text{SS} = 1 - |\braket{\Phi_0 | SS}|^2, \varepsilon_\text{CT} = 1 - |\braket{\Phi_0 | CT}|^2\). The spectral weight becomes
\begin{equation}\begin{aligned}
	d_n^p = \bra{\Phi_n} | c^\dagger_{d\sigma} \left(\ket{SS} + \ket{CT} \right) \left(\sqrt{1 - \varepsilon_\text{SS}} + \sqrt{1 - \varepsilon_\text{CT}}\right)
\end{aligned}\end{equation}
The total spectral weight can be written as
\begin{equation}\begin{aligned}
	|d_n^p|^2 = |\bra{\Phi_n} | c^\dagger_{d\sigma} \left(\ket{SS} + \ket{CT} \right)|^2 \left(2 - \varepsilon_\text{SS} - \varepsilon_\text{CT} + 2\sqrt{1 - \varepsilon_\text{SS}}\sqrt{1 - \varepsilon_\text{CT}}\right)
\end{aligned}\end{equation}
Near the metal, the ground state is, to a very good approximation, the state \(\ket{SS}\). One can then say that the entanglement \(\varepsilon_\text{SS}\) is poor - the entangled state is very close to the "separable" state. The term separable is used here in the sense that the metallic state defines a limit of the system. As \(U_b\) increases, \(\varepsilon_\text{SS}\) increases. On the other hand, the entanglement with the charge triplet is maximum in the metal, because the ground state does not have much of the charge triplet. Towards the MIT, this contribution increases, and the entanglement decreases.
